Geospatial - Assignment 2

### 2025-02-03 hw2/hw2_MoritzPeistNoemiLucchiSimonVellin.Rmd
updated RMD file

# Spatial packages
library(sf)
library(spData)
library(units)
```

@@

In order to plot the distribution of the total population by country, we only need the R packages world, which contains both the geometries of the countries and the corresponding population.

Given that we plot the population both in absolute terms and by categories — low, medium, and high — we need to define the population ranges for these categories. We classify countries with a population up to the 30th percentile as "Low population" countries, those with a population between the 30th and 70th percentiles as "Medium population" countries, and those in the top 30% of the population distribution as "High population" countries.

@@

# Divide the population by categories
world <- world %>%
  mutate(pop_category = case_when(
    pop <= quantile(pop, 0.30, na.rm = TRUE) ~ "Low",
    pop <= quantile(pop, 0.70, na.rm = TRUE) ~ "Medium",
    TRUE ~ "High"
  ))

@@

# Population in absolute terms
ggplot() +
  geom_sf(data = world, aes(fill = pop)) +
  scale_fill_gradient(name = "Population", low = "lightblue", high = "blue")

# Population by categories 
ggplot() +
  geom_sf(data = world, aes(fill = pop_category), alpha = 0.7) +
  scale_fill_manual(
    name = "Population Category",
    values = c("Low" = "lightyellow", "Medium" = "lightgreen", "High" = "blue"))

@@

world <- world %>%
  mutate(
    continent = case_when(
      continent %in% c("South America", "North America") ~ "America",
      continent %in% c("Seven seas (open ocean)", "Antarctica") ~ NA_character_,
      TRUE ~ continent
    )
  ) %>%
  filter(!is.na(continent))

@@

sf.airports <- st_read("hw2/data/ne_10m_airports", crs = 4326)
sf.pop <- st_read("hw2/data/ne_10m_populated_places_simple", crs = 4326)

>>>>

# Spatial packages
library(sf)
library(spData)
library(tidyverse)
library(units)
library(lwgeom)
```

@@

In order to plot the distribution of the total population by country, we import data from [Natural Earth](https://www.naturalearthdata.com), specifically the countries dataset which contains both the geometries of the countries and the corresponding population.

@@

sf.world <- st_read("data/ne_10m_admin_0_countries", crs = 4326)

@@

# Population in absolute terms
ggplot() +
  geom_sf(data = sf.world, aes(fill = POP_EST)) +
  scale_fill_gradient(name = "Population", low = "lightblue", high = "blue")

@@

# Remove Seven seas and Antarctica because they are NA
world <- world %>%
  mutate(
    continent = case_when(
      continent %in% c("Seven seas (open ocean)", "Antarctica") ~ NA_character_,
      TRUE ~ continent
    )
  ) %>%
  filter(!is.na(continent))

@@

sf.airports <- st_read("data/ne_10m_airports", crs = 4326)
sf.pop <- st_read("data/ne_10m_populated_places_simple", crs = 4326)
